---
title: "Rejection sampling"
author: Yan Shuo Tan
output: html_document
---

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
source("sampling_util.R")
```

In this demo, we show how to do rejection sampling for a density $f(x)$. We have access to a function $h(x) = cf(x)$ whose graph is plotted below and which is available to us as `myster_func`.

```{r Plot the mystery function}
ggplot(data.frame(x = c(0, 1)), aes(x = x)) + 
  stat_function(fun = mystery_func)
```

We will use the uniform distribution as the proposal $g(x)$. Hence, to compute $M$ so that $M g(x)$ such that $h(x) \leq M g(x)$, we simply need to compute the maximum of $h(x)$.

```{r Find the maximum of the function}
optim(0.5, fn = mystery_func, control = list(fnscale=-1))
```
We now run our script to do rejection sampling.

```{r Generate samples}
S <- 10000
M <- 4.99
proposal_generator <- function() {
  runif(1)
}
sampling_results <- rej_sampling_1d(S, target = mystery_func, proposal = dunif, 
                                    proposal_generator = proposal_generator, 
                                    M = M)
```

Compute an estimate $\hat c$ for the constant $c = \frac{f(x)}{h(x)}$.

```{r Estimate c}
c_hat <- sampling_results$accept_rate * M
c_hat
```

Finally, we plot a histogram of the samples we obtained, against the graph of rescaled $h(x)$.

```{r Plot results}
ggplot(data = tibble(x = sampling_results$samples)) + 
  geom_histogram(aes(x = x, y = ..density..), bins = 50) +
  stat_function(data = tibble(x = c(0, 1)), fun = function(x) {1 / c_hat * mystery_func(x)})
```
